---
title: "FiveThirtyEightGraphs"
output: html_document
---

```{r}
library(fivethirtyeight)
library(tidyverse)
```
```{r}
str(comic_characters)
head(comic_characters)
```

```{r}
ggplot(data = comic_characters, mapping = aes(x = id)) + geom_density()
```
This is a one variable graph. It is a density plot about the identity (id) of the comic book characters.

```{r}
ggplot(data = comic_characters, mapping = aes(x = id, y = month)) +geom_point()
```
A two variable graph that shows the relationship between the names of the comic book characters and their appearances.

```{r}
ggplot(data = comic_characters, mapping = aes(x = id, y = month, size = appearances)) +geom_point()
```
A three variable graph that shows the relation between character identity, the number of appearances, and the month of their appearance. The point size decrease as the number of appearances decrease.

```{r}
ggplot(data = comic_characters, mapping = aes(x = year, y = count(gsm))) +geom_bar(stat = "identity")
```

This graph is similar to last graph from the website. It shows the relationship between LGBT characters introduced per year. 